---
title: "Descriptives"
bibliography: references.bib
link-citations: true
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: false
    toc_depth: 2
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test3"))
options(width = 100)
rgl::setupKnitr()
options(knitr.kable.NA='')

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



---  
  
# Getting started

To copy the code, click the button in the upper right corner of the code-chunks.

## clean up

```{r, clean, results='hide'}
rm(list=ls())
gc()
```

<br>

## general custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R
- `fsave`: Function to save data with time stamp in correct directory
- `fload`: Load R-objects under new names
- `fshowdf`: Print objects (`tibble` / `data.frame`) nicely on screen in `.Rmd`.


```{r, functions}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}


fload <- function(fileName) {
    load(fileName)
    get(ls()[ls() != "fileName"])
}

fshowdf <- function(x, digits = 2, ...) {
    knitr::kable(x, digits = digits, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

#extrafont::font_import(paths = c("C:/Users/u244147/Downloads/Jost/", prompt = FALSE))
ftheme <- function() {
  
  #download font at https://fonts.google.com/specimen/Jost/
  theme_minimal(base_family = "Jost") +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(family = "Jost", face = "bold"),
          axis.title = element_text(family = "Jost Medium"),
          axis.title.x = element_text(hjust = 0),
          axis.title.y = element_text(hjust = 1),
          strip.text = element_text(family = "Jost", face = "bold",
                                    size = rel(0.75), hjust = 0),
          strip.background = element_rect(fill = "grey90", color = NA),
          legend.position = "bottom")
}
```

<br>

## necessary packages

- `tidyverse`
- `knitr`: generating tables
- `kableExtra`: manipulating tables
- `ergm.ego` 

```{r, packages, results='hide', message=FALSE, warning=FALSE}
packages = c("tidyverse", "knitr", "kableExtra", "ergm.ego")
fpackage.check(packages)
```

<br>

## load data

Load the replicated data-set (constructed [here](prep.html)). To load the file, adjust the filename in the following code so that it matches the most recent version of the `.RDa` file you have in your `./data/processed/` folder.

You may also obtain them by downloading: `r xfun::embed_file("./data_shared/egodata.RDa")`

```{r, data}
#get todays date:
today <- gsub("-", "", Sys.Date())

data <- fload(paste0("./data/processed/", today, "egodata.Rda"))
```


---

# Exploratory analysis 

## Mixing {.tabset .tabset-fade}

### Gender

```{r, exp, eval = TRUE}
layout(matrix(1:2, 1, 2))

barplot(table(data$ego$gender)/nrow(data$ego),
        main="Ego-gender distribution", ylab="proportion",
        ylim = c(0,.8), las = 3)
barplot(table(data$alter$gender)/nrow(data$alter),
        main="Alter-gender distribution", ylab="proportion",
        ylim = c(0,0.8), las = 3)

fshowdf(mixingmatrix(data, "gender"), caption = "Crosstabulated tie counts")
fshowdf(round(mixingmatrix(data, "gender", rowprob = TRUE), 2), caption = "Row conditional probabilities")
```

### LR-placement

```{r, exp2, eval = TRUE}
layout(matrix(1:2, 1, 2))

barplot(table(data$ego$LR)/nrow(data$ego),
        main="Ego LR-placement distribution", ylab="proportion",
        ylim = c(0,.25))
barplot(table(data$alter$LR)/nrow(data$alter),
        main="Alter LR-placement distribution", ylab="proportion",
        ylim = c(0,0.25))
```

## {.unlisted .unnumbered}

<br>

## Degrees

```{r, deg}
nrow(data$alter)/2
summary(data ~ degree(0:5))
degreedist(data, by = "gender", plot = TRUE, prob = FALSE)
degreedist(data, by = "gender", plot = TRUE, prob = TRUE)
degreedist(data, brg=TRUE)
degreedist(data, by="gender", prob=TRUE, brg=TRUE)
``` 

---

# Simple ERGM

```{r, ergm}
#edges only
fit.edges <- ergm.ego(data ~ edges)
summary(fit.edges)
fit.edges$ppopsize
fit.edges$popsize

summary(ergm.ego(data ~ edges, 
                 control = control.ergm.ego(ppopsize=10000)))
``` 
<br>

## Convergence checks

```{r, conv}
mcmc.diagnostics(fit.edges, which ="plots")
```

<br>

## GOF assessment

```{r, gof}
plot(gof(fit.edges, GOF="model"))
plot(gof(fit.edges, GOF="degree"))
```

```{r}
fitdeg5 <- ergm.ego(data ~ edges + degree(5))

summary(fitdeg5)
plot(gof(fitdeg5, GOF="degree"))
``` 
