---
title: "Data preparation"
bibliography: references.bib
link-citations: true
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: false
    toc_depth: 2
    code_folding: show
    code_download: yes
---


```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test3"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


---  

The following scripts can be used to replicate the data-set of @Franken2024. It may also be obtained by downloading: `r xfun::embed_file("./data_shared/egodata.RDa")`


----

# Getting started

To copy the code, click the button in the upper right corner of the code-chunks.

## clean up

```{r, cleanup, results='hide'}
rm(list=ls())
gc()
```

<br>

## general custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R
- `fsave`: Function to save data with time stamp in correct directory
- `fload`: Load R-objects under new names
- `fshowdf`: Print objects (`tibble` / `data.frame`) nicely on screen in `.Rmd`.

```{r, customfunc, eval=FALSE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}


fload  <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}

fshowdf <- function(x, digits = 2, ...) {
    knitr::kable(x, digits = digits, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```

<br>

## necessary packages

- `tidyverse` 
- `igraph` 
- `egor`

```{r, packages, eval=FALSE}
packages = c("tidyverse", "igraph", "egor")
fpackage.check(packages)
rm(packages)
```

----


# Download data

Anonymized data-sets of the 'Sports and Friendships' study are deposited in DANS Data Station SSH [@data2022]. To see the code used to anonymize the raw data files, see: https://netchange.netlify.app/prep. For this study, we use wave 3 (Cohort II):

- `r xfun::embed_file("./data/wave3_public.RDa")`

<br>

Download the data, and put the file in the `./data/` folder. But first, make a `./data/` folder: 

```{r, download, eval=F}
ifelse(!dir.exists("data"), dir.create("data"), FALSE)
``` 

----

# Import data

Load the downloaded data. First, we clean our environment (but we keep our functions; we need them later on).

```{r, import, eval = FALSE}
#clean environment, but keep functions
rm(list = setdiff(ls(), lsf.str()))

#load public data
data <- fload("./data/wave3_public.RDa")
```

---

# Data wrangling

```{r, wrangling, eval=FALSE, warning=FALSE}
#1. ego data-frame:
egodat <- data.frame(
  ego = 1:nrow(data), # ego_id
  gender = data$A1,   # ego gender
  LR = as.numeric(gsub("[^0-9]", "", data$P1)) # ego left-right placement
)

#2. list of data-frames of alters
altdat <- vector("list", nrow(egodat))

#we take alters from the "core discussion network" (i.e., egonet1)
cdn <- paste0("egonet1.SQ00", sprintf("%d", 1:5), ".")

#and, for now, their gender and LR-placement (according to ego!)
data %>%
  select(gender.SQ001.:gender.SQ005.) -> genders

data %>%
  select(P5.SQ001.:P5.SQ005.) %>%
  mutate_all(~ as.numeric(gsub("[^0-9.]", "", .))) -> LRs

#loop over all egos;
for (i in 1:length(altdat)) {
  
  #naturally, only for egos who listed at least 1 alter/confidant

  if(sum(data[i, cdn] != "") > 0) {
    
    altdat[[i]] <- data.frame(
      ego = rep(i, length(grep("p", data[i, cdn]))),
      alter = NA,
      gender = NA,
      LR = NA)

    #make unique alter-id
    altdat[[i]]$alter <- as.numeric(paste0(i, "0", grep("p", data[i, cdn])))
    
    #retrieve alters attributes
    altdat[[i]]$gender <- as.vector(unlist(genders[i, 1:nrow(altdat[[i]])]))
    altdat[[i]]$LR <- as.vector(unlist(LRs[i, 1:nrow(altdat[[i]])]))
  }
}

#3. list of edge lists (alter-alter ties, nested in ego)

aatie <- vector("list", nrow(egodat))

for (i in 1:length(aatie)) {
  
  #only for egos/list elements containing at least 2 alters:
  netsize <- sum(data[i, cdn] != "")
  
  if(netsize > 1) {
    
    #make list of adjacency matrices
    adjL <- list()
    
    {
      #2 alters
      adjL[[2]] <- matrix(NA,ncol=2,nrow=2)
      adjL[[2]][1,2] <- adjL[[2]][2,1] <- data$adj1N2a.SQ001.[i]

      #3 alters
      adjL[[3]] <- matrix(NA,ncol=3,nrow=3)
      adjL[[3]][1,2] <- adjL[[3]][2,1] <- data$adj1N3a.SQ001.[i]
      adjL[[3]][1,3] <- adjL[[3]][3,1] <- data$adj1N3a.SQ002.[i]
      adjL[[3]][2,3] <- adjL[[3]][3,2] <- data$adj1N3b.SQ001.[i]
    
      #4 alters
      adjL[[4]] <- matrix(NA,ncol=4,nrow=4)
      adjL[[4]][1,2] <- adjL[[4]][2,1] <- data$adj1N4a.SQ001.[i]
      adjL[[4]][1,3] <- adjL[[4]][3,1] <- data$adj1N4a.SQ002.[i]
      adjL[[4]][1,4] <- adjL[[4]][4,1] <- data$adj1N4a.SQ003.[i]
      adjL[[4]][2,3] <- adjL[[4]][3,2] <- data$adj1N4b.SQ001.[i]
      adjL[[4]][2,4] <- adjL[[4]][4,2] <- data$adj1N4b.SQ002.[i]
      adjL[[4]][3,4] <- adjL[[4]][4,3] <- data$adj1N4c.SQ001.[i]
    
      #5 alters
      adjL[[5]] <- matrix(NA,ncol=5,nrow=5)
      adjL[[5]][1,2] <- adjL[[5]][2,1] <- data$adj1N5a.SQ001.[i]
      adjL[[5]][1,3] <- adjL[[5]][3,1] <- data$adj1N5a.SQ002.[i]
      adjL[[5]][1,4] <- adjL[[5]][4,1] <- data$adj1N5a.SQ003.[i]
      adjL[[5]][1,5] <- adjL[[5]][5,1] <- data$adj1N5a.SQ004.[i]
      adjL[[5]][2,3] <- adjL[[5]][3,2] <- data$adj1N5b.SQ001.[i]
      adjL[[5]][2,4] <- adjL[[5]][4,2] <- data$adj1N5b.SQ002.[i]
      adjL[[5]][2,5] <- adjL[[5]][5,2] <- data$adj1N5b.SQ003.[i]
      adjL[[5]][3,4] <- adjL[[5]][4,3] <- data$adj1N5c.SQ001.[i]
      adjL[[5]][3,5] <- adjL[[5]][5,3] <- data$adj1N5c.SQ002.[i]
      adjL[[5]][4,5] <- adjL[[5]][5,4] <- data$adj1N5d.SQ001.[i]
    }
    
    #take the list entry (i.e., matrix) corresponding to the CDN/network size of ego i
    mat <- adjL[[netsize]]
    
    #alter-alter ties that are (very) close are 1; others are 0
    mat <- apply(mat, c(1,2), function(x) ifelse(x %in% c("Erg hecht", "Hecht"), 1, 0))
    
    #to edgelist (i.e., find indices where value is 1)
    edge_list <- which(mat == 1, arr.ind = TRUE)
    
    #modify values such that they reflect unique alter-ids of altdat
    
    if(nrow(edge_list) > 0) {
      edge_list <- matrix(paste0(i, "0", edge_list), ncol=2)
    }
    
    #attach ego id
    edge_list <- cbind(i, edge_list)
    colnames(edge_list) <- c("ego", "send", "rec")
    
    #put in list
    aatie[[i]] <- edge_list
  }
}

# convert data.frames totibble
e <- as_tibble(egodat)
a <- as_tibble(do.call(rbind, altdat))
aa <- as_tibble(do.call(rbind, aatie))

# convert to egor object
egodata <- egor(egos = e,
             alters = a,
             aaties = aa,
             ID.vars = list(ego = "ego", 
                            alter = "alter",
                            source = "send",
                            target = "rec"),
             alter_design = list(max = 5)
             )

#labeling/order gender variable
egodata$ego$gender[egodata$ego$gender == "Overige"] <- "Ander"
egodata$ego$gender <- factor(egodata$ego$gender, levels = c("Man", "Vrouw", "Ander"))
egodata$alter$gender <- factor(egodata$alter$gender, levels = c("Man", "Vrouw", "Ander"))
```

----

# Save data set

```{r, fsave, eval = FALSE}
fsave(egodata, "egodata.Rda")
```


----

### References

